Revisiting the Austral Spring Extratropical Southern Hemisphere zonally asymmetric circulation using complex Empirical Orthogonal Functions

### 2022-06-07 analysis/paper/shceof.Rmd
Pasa todo al nuevo template

---
title: "Revisiting the Austral Spring Extratropical Southern Hemisphere zonally asymmetric circulation using complex Empirical Orthogonal Functions"
titlerunning:  SH zonally asymmetric circulation
authors: 
- name: Elio Campitelli
  address: |
    Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmósfera y los Océanos. Buenos Aires, Argentina.  
    CONICET – Universidad de Buenos Aires. Centro de Investigaciones del Mar y la Atmósfera (CIMA). Buenos Aires, Argentina.  
    CNRS – IRD – CONICET – UBA. Instituto Franco-Argentino para el Estudio del Clima y sus Impactos (IRL 3351 IFAECI). Buenos Aires, Argentina.
  email: elio.campitelli@cima.fcen.uba.ar

- name:  Leandro B. Díaz
  address: |
    Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmósfera y los Océanos. Buenos Aires, Argentina.
    CONICET – Universidad de Buenos Aires. Centro de Investigaciones del Mar y la Atmósfera (CIMA). Buenos Aires, Argentina.
    CNRS – IRD – CONICET – UBA. Instituto Franco-Argentino para el Estudio del Clima y sus Impactos (IRL 3351 IFAECI). Buenos Aires, Argentina.

- name:  Carolina Vera
  address: |
    Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmósfera y los Océanos. Buenos Aires, Argentina.
    CONICET – Universidad de Buenos Aires. Centro de Investigaciones del Mar y la Atmósfera (CIMA). Buenos Aires, Argentina.
    CNRS – IRD – CONICET – UBA. Instituto Franco-Argentino para el Estudio del Clima y sus Impactos (IRL 3351 IFAECI). Buenos Aires, Argentina.

thanks: | 
  The research was supported by UBACyT20020170100428BA and the CLIMAX Project funded by Belmont Forum/ANR-15-JCL/-0002-01. Elio Campitelli was supported by a PhD grant from CONICET, Argentina.

@@

date: "`r format(Sys.time(), '%d %B, %Y')`"

@@

bibstyle: spbasic
# bibstyle options spbasic(default), spphys, spmpsci
output: 
  # bookdown::word_document2: default    
  bookdown::pdf_book: 
      base_format: rticles::springer_article
      latex_engine: "xelatex"
      keep_tex: true
      # citation_package: "natbib"
      pandoc_args: ["--lua-filter=abstract-to-meta.lua"]
always_allow_html: true
header-includes: 
 - \usepackage{gensymb}
 - \usepackage{subfig}
 - \usepackage[inline]{showlabels}
 - \usepackage{chngcntr}
 - \usepackage{natbib}
 - \usepackage{lineno}
 - \linenumbers 
---

@@

```{r extended-series, fig.cap = "(ref:extended-series-cap)", fig.height=3, fig.width = 6, fig.env = "figure*"}
# Check that the extended index coincides with the
# original index.
cors <- era_more %>%
  .[sep_ReIm(ceof$eof[[1]]$left), on = c("time", "cEOF", "part")] %>%
  na.omit() %>%
  .[, cor(hgt, i.hgt), by = .(cEOF, part)]

@@

era_more %>%
  copy() %>% 
  .[, ref := year(time) > 1979] %>% 
  # .[, hgt := scale(hgt, center = FALSE, scale = sd(hgt[ref])), by = .(part, cEOF)] %>%
  ggplot(aes(time, hgt*1e3)) +
  # Zero line + rangeframe
  geom_hline(data = ~.x[, mean(hgt)*1e3, by  = .(cEOF, part)], 
             aes(yintercept = V1), size = 0.2, color = "gray30") +
  geom_segment(data = ~.x[, .(min = min(hgt)*1e3,
                              max = max(hgt)*1e3), by  = .(cEOF, part)],
               aes(x = structure(Inf, class = "Date"),
                   xend = structure(Inf, class = "Date"),
                   y = min,
                   yend = max), size = 0.2, color = "gray30") +
  geom_line(aes(color = part)) +
  geom_point(aes(color = part), size = 0.25) +
  # geom_line(data = era_less) +
  scale_alpha_manual(values = c(1, 0.2)) +
  geom_smooth(method = "lm", aes(color = stage(part, after_scale = scales::muted(color))),
              show.legend = FALSE, size = 0.8) +
  scale_x_date(NULL, date_labels = "%Y",
               breaks = seq(as.Date("1950-01-01"),
                            as.Date("2020-01-01"), by = "10 years"),
               minor_breaks = seq(as.Date("1950-01-01"),
                                  as.Date("2020-01-01"), by = "2 years")) +
  scale_y_continuous(NULL, minor_breaks = NULL, sec.axis = dup_axis()) +
  scale_color_brewer(NULL, palette = "Dark2", 
                     guide = guide_legend(override.aes = list(fill = NA))) +
  facet_grid(cEOF ~ part) +
  theme(panel.spacing = grid::unit(1.5, "lines"),
        strip.placement = "outside", 
        panel.grid.minor.x = element_line(linetype = 2))  +
  tag_facets("rc") +
  panel_background 
```

@@

    facet_grid(angle ~ var, labeller = labeller(angle = angles_labs)) +
    # coord_quickmap(ylim = c(NA, 10)) +
    coord_sf(ylim = c(-89, 10), expand = FALSE,
             # crs = "+proj=longlat +pm=180 +over",
             default_crs = "+proj=longlat +lon_wrap=180 +over") + 
    theme(legend.box = "horizontal",
          legend.text = element_text(size = rel(0.7)), 
          legend.box.spacing = grid::unit(0, "lines") ) +
    tag_facets("rc") 
}
```

```{r  sst-psi-2,  fig.cap = "(ref:sst-psi-2-cap)", fig.width = 7, fig.height=5.6, dependson="plot_sst_psi", fig.env="figure*"}
plot_sst_psi(sst_regr[cEOF == "cEOF2"], psi_regr[cEOF == "cEOF2"]) 
```

@@

```{r pp-t2m-r2, fig.cap = "(ref:cap-pp-t2m-r2)", fig.width=3.3*2, fig.height=3.3}
breaks <- seq(.2, 1, by = .1)

@@

    scale_x_continuous("", expand = c(0, 0), breaks = seq(0, 360, by = 30), 
                       labels = LonLabel) +
    scale_y_continuous("", expand = c(0, 0), breaks = seq(-90, 0, by = 15), 
                       labels = LatLabel) +
    # scale_y_latitude(ticks = 15, labels = LatLabel) +
    scale_linetype_discrete(guide = "none") +
    facet_grid(angle ~ var, labeller = labeller(angle = angles_labs)) +
    axis_labs_smol +
    tag_facets("rc") +
    coord_sf(ylim = c(NA, 10),
             default_crs = "+proj=longlat +over") +
    theme(panel.spacing = grid::unit(1, "lines")) +
    theme(legend.box = "horizontal",
          legend.text = element_text(size = rel(0.7)), 
          legend.box.spacing = grid::unit(0, "lines") )

@@

```{r pp-temp-2, fig.cap="(ref:pp-temp-cap)", fig.width=6.6, fig.height=5.5, dependson="plot_pp"}
plot_pp(pp_regr[cEOF == "cEOF2"], t2m_regr[cEOF == "cEOF2"], hgt850_regr[cEOF == "cEOF2"])
```

@@

@@

```{r sst-psi-1, fig.cap = "(ref:sst-psi-1-cap)", fig.width = 7, fig.height=5.6, dependson="plot_sst_psi", fig.env = "figure*"}
plot_sst_psi(sst_regr[cEOF == "cEOF1"], psi_regr[cEOF == "cEOF1"]) 
```

>>>>

---
classoptions: 
  - sn-basic
year: 2022
title:  "Revisiting the Austral Spring Extratropical Southern Hemisphere zonally asymmetric circulation using complex Empirical Orthogonal Functions"
titlerunning: SH zonally asymmetric circulation with cEOF

authors: 
  - firstname: Elio
    lastname: Campitelli
    email: elio.campitelli@cima.fcen.uba.ar
    affiliation: [1,2,3]
    corresponding: TRUE
  - firstname: Leandro B.
    lastname: Díaz
    affiliation: [1,2,3]
  - firstname: Carolina
    lastname: Vera
    affiliation: [1,2,3]

affiliations:
  - number: 1
    corresponding: TRUE
    info:
      orgname: Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Ciencias de la Atmósfera y los Océanos
    address:
        city: Buenos Aires
        country: Argentina
  - number: 2
    corresponding: TRUE
    info:
      orgname: CONICET – Universidad de Buenos Aires. Centro de Investigaciones del Mar y la Atmósfera (CIMA)
    address:
        city: Buenos Aires
        country: Argentina  
  - number: 3
    corresponding: TRUE
    info:
      orgname: CNRS – IRD – CONICET – UBA. Instituto Franco-Argentino para el Estudio del Clima y sus Impactos (IRL 3351 IFAECI)
    address:
        city: Buenos Aires
        country: Argentina

@@

@@

# bibstyle: spbasic

output: 
  bookdown::pdf_document2:
      latex_engine: "xelatex"
      keep_tex: true
      citation_package: "natbib"
      template: "template.tex"
      pandoc_args: ["--lua-filter=abstract-to-meta.lua", "--wrap=preserve"]
always_allow_html: true
header-includes: 
 - \usepackage{gensymb}
 - \usepackage{subfig}
 - \usepackage[inline]{showlabels}
 - \usepackage{chngcntr}
 # - \usepackage{natbib}
 # - \usepackage{lineno}
 # - \linenumbers 
---

@@

```{r extended-series, fig.cap = "(ref:extended-series-cap)", fig.height=3, fig.width = 5, fig.env = "figure*"}
# Check that the extended index coincides with the
# original index.
cors <- era_more %>%
  .[sep_ReIm(ceof$eof[[1]]$left), on = c("time", "cEOF", "part")] %>%
  na.omit() %>%
  .[, cor(hgt, i.hgt), by = .(cEOF, part)]

@@

era_more %>%
  copy() %>% 
  .[, ref := year(time) > 1979] %>% 
  # .[, hgt := scale(hgt, center = FALSE, scale = sd(hgt[ref])), by = .(part, cEOF)] %>%
  ggplot(aes(time, hgt*1e3)) +
  # Zero line + rangeframe
  geom_hline(data = ~.x[, mean(hgt)*1e3, by  = .(cEOF, part)], 
             aes(yintercept = V1), size = 0.2, color = "gray30") +
  geom_segment(data = ~.x[, .(min = min(hgt)*1e3,
                              max = max(hgt)*1e3), by  = .(cEOF, part)],
               aes(x = structure(Inf, class = "Date"),
                   xend = structure(Inf, class = "Date"),
                   y = min,
                   yend = max), size = 0.2, color = "gray30") +
  geom_line(aes(color = part)) +
  geom_point(aes(color = part), size = 0.25) +
  # geom_line(data = era_less) +
  scale_alpha_manual(values = c(1, 0.2)) +
  geom_smooth(method = "lm", aes(color = stage(part, after_scale = scales::muted(color))),
              show.legend = FALSE, size = 0.8) +
  scale_x_date(NULL, 
               breaks = seq(as.Date("1950-01-01"),
                            as.Date("2020-01-01"), by = "10 years"),
               labels = function(x) JumpBy(format(x, "%Y"), 2, fill = " "),
               minor_breaks = seq(as.Date("1950-01-01"),
                                  as.Date("2020-01-01"), by = "2 years")) +
  scale_y_continuous(NULL, minor_breaks = NULL, sec.axis = dup_axis()) +
  scale_color_brewer(NULL, palette = "Dark2", 
                     guide = guide_legend(override.aes = list(fill = NA))) +
  facet_grid(cEOF ~ part) +
  theme(panel.spacing = grid::unit(1.5, "lines"),
        strip.placement = "outside", 
        panel.grid.minor.x = element_line(linetype = 2))  +
  tag_facets("rc") +
  panel_background 
```

@@

    facet_grid(angle ~ var, labeller = labeller(angle = angles_labs)) +
    # coord_quickmap(ylim = c(NA, 10)) +
    coord_sf(ylim = c(-89, 10), expand = FALSE,
             # crs = "+proj=longlat +pm=180 +over",
             default_crs = "+proj=longlat +lon_wrap=180 +over") + 
    theme(legend.box = "vertical",
          legend.text = element_text(size = rel(0.7)), 
          legend.spacing = grid::unit(0, "lines"),
          legend.box.spacing = grid::unit(0, "lines") ) +
    tag_facets("rc") 
}
```

```{r sst-psi-2,  fig.cap = "(ref:sst-psi-2-cap)", fig.width = 5, fig.height=6, dependson="plot_sst_psi", fig.env="figure*"}
plot_sst_psi(sst_regr[cEOF == "cEOF2"], psi_regr[cEOF == "cEOF2"]) 
```

@@

```{r pp-t2m-r2, fig.cap = "(ref:cap-pp-t2m-r2)", fig.width=5, fig.height=3}
breaks <- seq(.2, 1, by = .1)

@@

    scale_x_continuous("", expand = c(0, 0), breaks = seq(0, 360, by = 30), 
                       labels = LonLabel) +
    scale_y_continuous("", expand = c(0, 0), breaks = seq(-90, 0, by = 15), 
                       labels = LatLabel) +
    # scale_y_latitude(ticks = 15, labels = LatLabel) +
    scale_linetype_discrete(guide = "none") +
    facet_grid(angle ~ var, labeller = labeller(angle = angles_labs)) +
    axis_labs_smol +
    tag_facets("rc") +
    coord_sf(ylim = c(NA, 10),
             default_crs = "+proj=longlat +over") +
    theme(panel.spacing = grid::unit(1, "lines")) +
    theme(legend.box = "vertical",
          legend.spacing = grid::unit(0, "lines"),
          legend.text = element_text(size = rel(0.7)), 
          legend.box.spacing = grid::unit(0, "lines") )

@@

```{r pp-temp-2, fig.cap="(ref:pp-temp-cap)", fig.width=5, fig.height=6, dependson="plot_pp"}
plot_pp(pp_regr[cEOF == "cEOF2"], t2m_regr[cEOF == "cEOF2"], hgt850_regr[cEOF == "cEOF2"])
```

@@

\backmatter

\bmhead{Acknowledgments}

The research was supported by UBACyT20020170100428BA and the CLIMAX Project funded by Belmont Forum/ANR-15-JCL/-0002-01. Elio Campitelli was supported by a PhD grant from CONICET, Argentina.

@@

```{r sst-psi-1, fig.cap = "(ref:sst-psi-1-cap)", fig.width = 5, fig.height=6, dependson="plot_sst_psi", fig.env = "figure*"}
plot_sst_psi(sst_regr[cEOF == "cEOF1"], psi_regr[cEOF == "cEOF1"]) 
```
